Expected Dependency Pair Match:\\ Predicting translation quality with expected syntactic structure

### 2009-05-16 mtjournal.tex
Matt's edits

Machine translation (MT) evaluation is a challenge for research
because the space of good translations is large, and two equally good
translations may appear to be quite different at first glance. 
%
The challenges of choosing among translations are compounded when this
evaluation is done automatically.
%
Human evaluation, however, is both time-consuming and difficult, so
research has turned increasingly towards automatic measures of
translation quality, usually by comparing the system translation to
one or more reference (human) translations.
%
Automatic measures of this kind not
only provide a well-defined evaluation standard but are also required
for training on error criteria, e.g.\ with minimum error rate training
\cite{och03mert}.

@@

These challenges have motivated a search for better measures that
incorporate additional language knowledge sources.  METEOR
\cite{banerjee05meteor}, for example, uses synonym tables and
morphological stemming to do progressively more-forgiving matching.
It can be tuned towards recall or precision, but is generally not
tuned by users.  TERp \cite{snover09terp} is an extension of the
previously-mentioned TER that also incorporates synonym sets, along
with automatically-derived paraphrase tables.  TERp is explicitly
intended to be tuned to a development set by users.
%
Tuning has the advantage that the weight of different types of errors
can be adjusted to match the needs of the task, though it makes it
more difficult to compare results across tasks, particularly when
there is little data for tuning.

@@

In section~\ref{sec:approach}), we outline our approach to including
syntactic information in MT evaluation, by describing a family of
metrics. In section~\ref{sec:paradigm}, we describe the implementation
details and make explicit which dimensions of this family of metrics
we explore in this paper.  Section~\ref{sec:faexpts} examines the
correlation of several members of this family with human judgments of
fluency and adequacy over a corpus of sentence translations and
selects a favorable member of the family, Expected Dependency Pair
Match (EDPM), based on this comparison. Section~\ref{sec:hter1}
compares the correlation of EDPM with HTER over a corpus of document-
and sentence-level comparisons, and demonstrates that its correlation
is favorable to that of competitor measures BLEU and TER at both
sentence and document levels, regardless of source-language or genre.
Section~\ref{sec:hter2} uses the tuning components of the TERp tools
to explore the combination of these syntactic measures with other TERp
word-based measures. Finally, section~\ref{sec:conclusion} concludes
with some remarks on challenges and future work.

@@

A member of this family is defined by several parameters.  The first
of these parameters is the nature of the decomposition of the
dependency tree structure.  A \emph{decomposition list} is the list of
ways in which the tree is reduced to a bag of tree-local
tuples. Figure~\ref{fig:decompexample}
\begin{figure}
  \centering
  \begin{tabular}{rcc}
    & \textbf{Reference} & \textbf{Hypothesis} \\
    \textbf{tree}
    & \includegraphics[scale=0.5]{dpm-example-ref.eps} & 
    \includegraphics[scale=0.5]{dpm-example-hyp.eps}\\
    \textbf{$dlh$ list} &
    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
      the & \arclabel{\texttt{det}}  &  cat \\
      red & \arclabel{\texttt{mod}}  &  cat \\
      cat & \arclabel{\texttt{subj}} &  ate \\
      ate & \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular} & 
    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
      the &      \arclabel{\texttt{det}}  &  cat \\
      cat &      \arclabel{\texttt{subj}} &  stumbled \\
      stumbled & \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular}\\
  \end{tabular}\\
  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
  \caption{Example hypothesis and reference dependency trees and the
    $dlh$ decomposition of each.}
  \label{fig:decompexample}
\end{figure}
illustrates the \emph{dependency-link-head} decomposition of a toy
dependency tree into a list of $\langle d, l, h \rangle$ tuples.  Some
members of the DPM family may apply more than one decomposition; other
good examples are the $dl$ decomposition, which generates a bag of
dependent word with outbound links, and the $lh$ decomposition, which
generates a bag of inbound link labels, with the head word for each
included. 
\begin{figure}
  \centering
    \begin{tabular}{cc}
    $dl$ & $lh$ \\
    \hline
    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
      the & \arclabel{\texttt{det}}  \\
      cat & \arclabel{\texttt{subj}} \\
      stumbled & \arclabel{\texttt{root}} \\
    \end{tabular} &
    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
      \arclabel{\texttt{det} } &  cat \\
      \arclabel{\texttt{subj}} &  stumbled \\
      \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular}\\
  \end{tabular}
  \caption{The $dl$ and $lh$ decompositions of the hypothesis tree in
    figure~\ref{fig:decompexample}. The items extracted here
    are individually less restrictive in their ability to match
    against reference tuples. 
%     $\precision{dl} = \frac{2}{3}$,
%     $\recall{dl} = \frac{2}{4}$, $\precision{lh} = \frac{2}{3}$,
%     $\recall{lh} = \frac{2}{4}$, so $F[dl,lh] = \frac{4}{7}$
  }
  \label{fig:decompexample2}
\end{figure}
Figure~\ref{fig:decompexample2} demonstrates the $dl,lh$
decomposition for the hypothesis tree in figure~\ref{fig:decompexample}.

@@

In this section, we have presented experiments exploring a number of
variants of the DPM metric against an average fluency/adequacy
judgment. The experiments suggest a best-case variant, where we set:
\begin{displaymath}
  EDPM = F[1g,2g,dl,lh], n=50, \gamma=0.25
\end{displaymath}
in which we choose a $1g,2g,dl,lh$ sub-graph decomposition based on
the improvements from better sub-graphs
(table~\ref{tab:facorr:subgraphs}), multiple parses ($n=50$) based on
table~\ref{tab:facorr:multiparse}, and $\gamma=0.25$.
%, hinted at by table~\ref{tab:facorr:parseweight}.  
We use these EDPM parameter-settings in the experiments exploring
document-level correlations with HTER (section~\ref{sec:hter1}).

@@

%% MO: important to describe basic model before listing features.
The TERp optimizer performs a hill-climbing search, with randomized
restarts, to maximize the correlation of a linear combination of the
features with a set of human judgments.  Within the TERp framework the
features are the count of the various edit types normalized for the
length of the reference so that the dot product of the features with a
vector of edit costs will give the TERp score for a given segment.  In
these experiments, we follow~\inlinecite{snover09terp} by first
aligning the MT outputs with the reference, using TERp with a set of
default edit costs, to find the counts of the various edit types.
These alignments and edit costs are then held constant to optimize the
edit costs. In the case of comparing TERp without the paraphrasing
feature, a separate alignment was performed without that feature
enabled.  Additional features can be added to the TERp optimization
framework and optimized in the same manner as the edit costs.

In the experiments presented in this section, we extend this feature
list by including the syntactic overlap features, and we use the TERp
optimizer tools to tune the relative weight of each feature.

@@

We train a model using features from one or more of the following
sets:
\begin{description}
\item[E]: the fully syntactic features from the DPM family.
  Specifically, this feature set includes error-counts
  ($\frac{|\textrm{hyp}| - |\textrm{correct}|}{|\textrm{ref}|}$ and
  $\frac{|\textrm{ref}| - |\textrm{correct}|}{|\textrm{ref}|}$) 
  for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic
  structure, using $n=50$ and $\gamma=0.25$.  The \textbf{E} set
  includes 8 features.
\item[N]: the non-syntactic features from the DPM family;
  specifically, error-counts for the $1g$ and $2g$ decompositions.
  The \textbf{N} set includes 4 features.
\item[T]: the features from basic TERp, excluding paraphrases, which include a separate cost
  for each of matches, insertions, deletions, substitutions, shifts, and
  synonym matches and stem matches. The \textbf{T} set includes 7 features.
\item[P]: features from the automatically-derived TERp paraphrase
  table. These are four features which are a function of of the edit
  distance between the two phrases in each paraphrase and the
  probability of the paraphrase.
\item[B]: two ``length penalties'' one is assigned the count of
  hypothesis words beyond the count of reference words, and one
  assigned the count of reference words beyond the count of hypothesis
  wordsbrevity/prolixity (2 features: one for longer-than-ref, one for
  shorter-than-ref).
  % MO: out of curiosity, why do you need 2 features?
  We assign two features here to let the system separately penalize
  ``too short'' translations and ``too long''.
\end{description}

>>>>

Machine translation (MT) evaluation is a challenge for research
because the space of good translations is large, and two equally good
translations may appear to be quite different at first glance. 
%
The challenges of choosing among translations are compounded when this
evaluation is done automatically.
%
Human evaluation, however, is both time-consuming and difficult, so
research has turned increasingly towards automatic measures of
translation quality, usually by comparing the system (hypothesis) translation to
one or more reference (human) translations.
%
Automatic measures of this kind not
only provide a well-defined evaluation standard but are also required
for training on error criteria, e.g.\ with minimum error rate training
\cite{och03mert}.

@@

These challenges have motivated a search for better measures that
incorporate additional language knowledge sources.  METEOR
\cite{banerjee05meteor}, for example, uses synonym tables and
morphological stemming to do progressively more-forgiving matching.
It can be tuned towards recall or precision, but is generally not
tuned by users.  TERp \cite{snover09terp} is an extension of the
previously-mentioned TER that also incorporates synonym sets and stemming, along
with automatically-derived paraphrase tables.  TERp is explicitly
intended to be tuned to a development set by users.
%
Tuning has the advantage that the weight of different types of errors
can be adjusted to match the needs of the task, though it makes it
more difficult to compare results across tasks, particularly when
there is little data for tuning.

@@

In section~\ref{sec:approach}, we outline our approach to including
syntactic information in MT evaluation, by describing a family of
metrics. In section~\ref{sec:paradigm}, we describe the implementation
details and make explicit which dimensions of this family of metrics
we explore in this paper.  Section~\ref{sec:faexpts} examines the
correlation of several members of this family with human judgments of
fluency and adequacy over a corpus of sentence translations and
selects a favorable member of the family, Expected Dependency Pair
Match (EDPM), based on this comparison. Section~\ref{sec:hter1}
compares the correlation of EDPM with HTER over a corpus of document-
and sentence-level comparisons, and demonstrates that its correlation
is favorable to that of competitor measures BLEU and TER at both
sentence and document levels, regardless of source-language or genre.
Section~\ref{sec:hter2} uses the tuning components of TERp
to explore the combination of these syntactic measures with TERp's
word-based measures. Finally, section~\ref{sec:conclusion} concludes
with some remarks on challenges and future work.

@@

A member of this family is defined by several parameters.  The first
of these parameters is the nature of the decomposition of the
dependency tree structure.  A \emph{decomposition list} is the list of
ways in which the tree is reduced to a bag of tree-local
tuples. Figure~\ref{fig:decompexample}
\begin{figure}
  \centering
  \begin{tabular}{rcc}
    & \textbf{Reference} & \textbf{Hypothesis} \\
    \textbf{tree}
    & \includegraphics[scale=0.5]{dpm-example-ref.eps} & 
    \includegraphics[scale=0.5]{dpm-example-hyp.eps}\\
    \textbf{$dlh$ list} &
    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
      the & \arclabel{\texttt{det}}  &  cat \\
      red & \arclabel{\texttt{mod}}  &  cat \\
      cat & \arclabel{\texttt{subj}} &  ate \\
      ate & \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular} & 
    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
      the &      \arclabel{\texttt{det}}  &  cat \\
      cat &      \arclabel{\texttt{subj}} &  stumbled \\
      stumbled & \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular}\\
  \end{tabular}\\
  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
  \caption{Example hypothesis and reference dependency trees and the
    $dlh$ decomposition of each.}
  \label{fig:decompexample}
\end{figure}
illustrates the \emph{dependency-link-head} decomposition of a toy
dependency tree into a list of $\langle d, l, h \rangle$ tuples.  Some
members of the DPM family may apply more than one decomposition; other
good examples are the $dl$ decomposition, which generates a bag of
dependent word with outbound links, and the $lh$ decomposition, which
generates a bag of inbound link labels, with the head word for each
included. 
\begin{figure}
  \centering
    \begin{tabular}{cc}
    $dl$ & $lh$ \\
    \hline
    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
      the & \arclabel{\texttt{det}}  \\
      cat & \arclabel{\texttt{subj}} \\
      stumbled & \arclabel{\texttt{root}} \\
    \end{tabular} &
    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
      \arclabel{\texttt{det} } &  cat \\
      \arclabel{\texttt{subj}} &  stumbled \\
      \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular}\\
  \end{tabular}
  \caption{The $dl$ and $lh$ decompositions of the hypothesis tree in
    figure~\ref{fig:decompexample}. The items extracted here
    are individually less restrictive in their ability to match
    against reference tuples. 
%     $\precision{dl} = \frac{2}{3}$,
%     $\recall{dl} = \frac{2}{4}$, $\precision{lh} = \frac{2}{3}$,
%     $\recall{lh} = \frac{2}{4}$, so $F[dl,lh] = \frac{4}{7}$
  }
  \label{fig:decompexample2}
\end{figure}
Figure~\ref{fig:decompexample2} demonstrates the $dl$ and $lh$
decompositions for the hypothesis tree in figure~\ref{fig:decompexample}.

@@

In this section, we have presented experiments exploring a number of
variants of the DPM metric against an average fluency/adequacy
judgment. The experiments suggest a best-case variant, where we set:
\begin{displaymath}
  \mbox{EDPM} = F[1g,2g,dl,lh], n=50, \gamma=0.25
\end{displaymath}
in which we choose a $1g,2g,dl,lh$ sub-graph decomposition based on
the improvements from better sub-graphs
(table~\ref{tab:facorr:subgraphs}), multiple parses ($n=50$) based on
table~\ref{tab:facorr:multiparse}, and $\gamma=0.25$.
%, hinted at by table~\ref{tab:facorr:parseweight}.  
We use these EDPM parameter-settings in the experiments exploring
document-level correlations with HTER (section~\ref{sec:hter1}).

@@

%% MO: important to describe basic model before listing features.
The TERp optimizer performs a hill-climbing search, with randomized
restarts, to maximize the correlation of a linear combination of the
features with a set of human judgments.  Within the TERp framework the
features are the counts of the various edit types normalized for the
length of the reference so that the dot product of the features with a
vector of edit costs will give the TERp score for a given sentence.  In
these experiments, we follow~\inlinecite{snover09terp} by first
aligning the MT outputs with the reference, using TERp with a set of
default edit costs, to find the counts of the various edit types.
These alignments and edit costs are then held constant to optimize the
edit costs. In the case of comparing TERp without the paraphrasing
feature, a separate alignment was performed without that feature
enabled.  Additional features can be added to the TERp optimization
framework and optimized in the same manner as the edit costs.

In the experiments presented in this section, we extend this feature
list by including the syntactic overlap features, and we use the TERp
optimizer to tune the relative weight of each feature.

@@

We train a model using features from one or more of the following
sets:
\begin{description}
\item[E]: the fully syntactic features from the DPM family.
  Specifically, this feature set includes error-counts
  ($\frac{|\textrm{hyp}| - |\textrm{correct}|}{|\textrm{ref}|}$ and
  $\frac{|\textrm{ref}| - |\textrm{correct}|}{|\textrm{ref}|}$) 
  for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic
  structure, using $n=50$ and $\gamma=0.25$.  The \textbf{E} set
  includes 8 features.
\item[N]: the non-syntactic features from the DPM family;
  specifically, error-counts for the $1g$ and $2g$ decompositions.
  The \textbf{N} set includes 4 features.
\item[T]: the features from basic TERp, excluding paraphrases, which include a separate cost
  for each of matches, insertions, deletions, substitutions, shifts, and
  synonym matches and stem matches. The \textbf{T} set includes 7 features.
\item[P]: features from the automatically-derived TERp paraphrase
  table. These are four features which are a function of the edit
  distance between the two phrases in each paraphrase and the
  probability of the paraphrase.
\item[B]: two ``length penalties'': brevity and prolixity. One is assigned the count of
  hypothesis words beyond the count of reference words, and one
  assigned the count of reference words beyond the count of hypothesis
  words.
  % MO: out of curiosity, why do you need 2 features?
  We assign two features here to let the system separately penalize
  ``too short'' translations and ``too long''.
\end{description}
